Visualising CHD Pathway Inequalities by ethnicity - DRAFT

### 2024-03-13 draft_report.qmd
draft report

This report quantifies and illustrates levels of inequity across 35 metrics at various points along the continuum of coronary heart disease progression and over a typical treatment pathway. They are shown in the table below grouped by domain (risk factors, risk factor identification, primary prevention, disease identification, secondary prevention, tertiary prevention, intermediate and full outcomes), which represent the various stages along the pathway. Full definitions and data sources for each metric are included in the appendix.

@@

...to be determined - awaiting death data check
:::

@@

-   The greatest disparity is flu vaccinations for patients aged 65+ where the relative index of disparity is 16.70% (95% ci - 16.64%, 16.70%), indicating that 16.7% of activity needs to be redistributed so that rates reflect need.
:::

@@

A previous Strategy Unit report "Strategies to reduce inequalities in access to planned hospital procedures"[^4] highlights in chapter 1, three of the many routes from inequity to equity: levelling-up, levelling-down and zero-sum redistribution. Using the zero-sum redistribution route as an example, along with the metric flu vaccinations for patients aged 65+, this illustration shows how equity could be achieved by increasing and decreasing the number of vaccinations delivered in each cluster to give the same activity rate per cluster.

@@

: In reality, in attempting to achieve equity for this metric, it is unlikely that the offer of vaccination would be removed from patients whose practice is in clusters 1 and 2. More likely would be an approach to focus public health and vaccination roll-out campaigns on practices in clusters 3, 4 and 5.

@@

-   The greatest disparity
:::

@@

.......

@@

| Domain                     | Metric                                 | Data Source | Year    | Definition and selection criteria/codes                                                 |
|---------------|---------------|---------------|---------------|---------------|
| Need                       | CHD synthetic prevalence estimates     | PHE         | 2015    | Provided directly by OHID (formerly indicator 92847 in FingertipsR)                     |
| Risk factors               | Smoking synthetic prevalence estimates | GP survey   | 2023    |                                                                                         |
| Risk factor identification | Smoking register                       | NHSD QOF    | 2022/23 | [Indicator 91280 extracted fromFingertipsR](https://fingertips.phe.org.uk/search/91280) |
|                            | Obesity register                       | NHSD QOF    |         | [Indicator 93088 extracted fromFingertipsR](https://fingertips.phe.org.uk/search/93088) |
|                            | Diabetes register                      | NHSD QOF    |         |                                                                                         |
|                            | Depression register                    | NHSD QOF    |         |                                                                                         |
| Primary prevention         |                                        |             |         |                                                                                         |
|                            |                                        |             |         |                                                                                         |
|                            |                                        |             |         |                                                                                         |
|                            |                                        |             |         |                                                                                         |
|                            |                                        |             |         |                                                                                         |
|                            |                                        |             |         |                                                                                         |
|                            |                                        |             |         |                                                                                         |
|                            |                                        |             |         |                                                                                         |
|                            |                                        |             |         |                                                                                         |
|                            |                                        |             |         |                                                                                         |
|                            |                                        |             |         |                                                                                         |
|                            |                                        |             |         |                                                                                         |
|                            |                                        |             |         |                                                                                         |
|                            |                                        |             |         |                                                                                         |
|                            |                                        |             |         |                                                                                         |
|                            |                                        |             |         |                                                                                         |
|                            |                                        |             |         |                                                                                         |
|                            |                                        |             |         |                                                                                         |
|                            |                                        |             |         |                                                                                         |
|                            |                                        |             |         |                                                                                         |
|                            |                                        |             |         |                                                                                         |
|                            |                                        |             |         |                                                                                         |
|                            |                                        |             |         |                                                                                         |
|                            |                                        |             |         |                                                                                         |
| Full Outcomes              | Deaths in hospital from CHD            |             |         |                                                                                         |
|                            | Deaths in hospital from CHD \< 75      |             |         |                                                                                         |
|                            | Deaths from CHD                        |             |         |                                                                                         |
|                            | Deaths from CHD \<75                   |             |         |                                                                                         |

@@

>>>>

This report quantifies and illustrates levels of inequity across 33 metrics at various points along the continuum of coronary heart disease progression and over a typical treatment pathway. They are shown in the table below grouped by domain (risk factors, risk factor identification, primary prevention, disease identification, secondary prevention, tertiary prevention, intermediate and full outcomes), which represent the various stages along the pathway. Full definitions and data sources for each metric are included in the appendix.

@@

...to be determined - awaiting death data checks
:::

@@

-   The greatest disparity along the CHD pathway is flu vaccinations for patients aged 65+ where the relative index of disparity is 16.70% (95% ci - 16.64%, 16.70%), indicating that 16.7% of activity needs to be redistributed so that rates reflect need.

-   The most equitably distributed metric is blood pressure checks within the last 5 years for patients aged 45+ with a relative index of disparity of 0.8% (95% ci - 0.82%, 0.78%).

-   There are 7 metrics where the index of disparity is less than 5%, as well as blood pressure checks, these include metrics such as referrals to outpatient cardiology, outpatient DNAs and waiting times for elective procedures.

-   Out of the 33 metrics along the CHD pathway, 10 have an index of disparity of greater than 10%.

-   In general (but not exclusively) these greater levels of disparity occur in the disease identification and secondary prevention pathway domains and consequently the outcome metrics related to deaths from CHD.
:::

@@

A previous Strategy Unit report "Strategies to reduce inequalities in access to planned hospital procedures"[^4] highlights in chapter 1, three of the many routes from inequity to equity: levelling-up, levelling-down and zero-sum redistribution. Using the zero-sum redistribution route as an example, along with the metric flu vaccinations for patients aged 65+, the illustration below shows how equity could be achieved by increasing and decreasing the number of vaccinations delivered in each cluster to give the same activity rate per cluster.

@@

In reality, in attempting to achieve equity for this metric, it is unlikely that the offer of vaccination would be removed from patients whose practice is in clusters 1 and 2. More likely would be an approach to focus public health and vaccination roll-out campaigns on practices in clusters 3, 4 and 5.

@@

-   Looking at a regional level, the greatest disparity is in CT angiography in the South West, Midland, and the North West.
:::

@@

To be added as part of finalisation of the report.

@@

...

@@

The charts below are equivalent charts to those in the regional analysis section of the report, showing the Index of Disparity for each CHD pathway metric for each of the 42 ICBs. There is considerable variation in the levels of disparity across the ICBs as well as the metrics along the pathway.
